Abstraction and Invariance for Algebraically Indexed Types

### 2012-07-09 paper/geometry-twocolumn.tex
Add caveats about quantifer-free ness

\begin{example}\label{ex:uninhabited-type}
  So far we have emphasised the derivation of properties, or ``free
  theorems'', of programs from their types. But using more refined
  relational interpretations of types we can also show that certain
  types are uninhabited. For example, the type $\forall t \mathord:
  \SynTransl{2}.\ \tyPrim{vec}{t + t} \to \tyPrim{vec}{t}$ has no
  inhabitants. Intuitively, this is because there is no way to remove
  the extra $t$ in $\tyPrim{vec}{t+t}$ using the vector operations. We
  formalise this indefinability result in
  \autoref{sec:types-indexed-abelian-groups} using a more
  sophisticated relational interpretation of free translation
  variables than the one we sketched above.
\end{example}

@@

\begin{example}
  With the operations in \autoref{fig:real-ops}, it is not possible to
  write a program with the following type that is not constantly zero:
  \begin{displaymath}
    \forall s \mathord: \SynGL{2}.\ \tyPrim{real}{s^2} \to \tyPrim{real}{s}
  \end{displaymath}
  This was shown by Kennedy for his units of measure system
  \cite{kennedy97relational}.  In particular, it is not possible to
  write a square root function with the above type.  The
  indefinability of square root is similar to the uninhabitation of the
  type in \exref{ex:uninhabited-type}.

  In \autoref{sec:types-indexed-abelian-groups} we revisit Kennedy's
  proof and show that even if we add square root as a primitive
  operation---with the type above---then it is still not possible to
  construct the cube root function. The indefinability of cube root is
  related to the impossibility of trisecting an arbitrary angle by
  ruler and compass constructions.
%  is of interest due to its relevance to the classical problem of
%  trisecting an angle by ruler and compass constructions.
\end{example}

>>>>

\begin{example}\label{ex:uninhabited-type}
  So far we have emphasised the derivation of properties, or ``free
  theorems'', of programs from their types. But using more refined
  relational interpretations of types we can also show that certain
  types are uninhabited. For example, the type $\forall t \mathord:
  \SynTransl{2}.\ \tyPrim{vec}{t + t} \to \tyPrim{vec}{t}$ has no
  quantifier-free inhabitants\footnote{We define quantifier-free
    programs in \thmref{thm:abstraction}.}. Intuitively, this is
  because there is no way to remove the extra $t$ in
  $\tyPrim{vec}{t+t}$ using the vector operations. We formalise this
  indefinability result in \autoref{sec:types-indexed-abelian-groups-indef}
  using a more sophisticated relational interpretation of free
  translation variables than the one we sketched above.
\end{example}

@@

\begin{example}
  With the operations in \autoref{fig:real-ops}, it is not possible to
  write a quantifier-free program with the following type that is not
  constantly zero:
  \begin{displaymath}
    \forall s \mathord: \SynGL{2}.\ \tyPrim{real}{s^2} \to \tyPrim{real}{s}
  \end{displaymath}
  This was shown by Kennedy for his units of measure system
  \cite{kennedy97relational}.  In particular, it is not possible to
  write a square root function with the above type.  The
  indefinability of square root is similar to the uninhabitation of the
  type in \exref{ex:uninhabited-type}.

  In \autoref{sec:types-indexed-abelian-groups-indef} we revisit
  Kennedy's proof and show that even if we add square root as a
  primitive operation---with the type above---then it is still not
  possible to construct the cube root function. The indefinability of
  cube root is related to the impossibility of trisecting an arbitrary
  angle by ruler and compass constructions.
%  is of interest due to its relevance to the classical problem of
%  trisecting an angle by ruler and compass constructions.
\end{example}
